---
title: "DD HS"
author: "LBE"
date: "23/05/2022"
output: github_document
---


#workspace setup

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(ggpubr)
library(aod)

```
#THEMES
```{r}
# paper theme
theme_DDHS <- theme_classic()+
    theme(panel.border = element_rect(size=1, fill=NA),
          axis.text = element_text(family="Arial", size = 12),
          axis.title = element_text(family="Arial", size = 12),
          title = element_text(family = "Arial",  size = 12),
          plot.title = element_text(hjust = 0.5, vjust = 1))

 cols_= c("#F7444E", "#78BCC4")
```


#SURVEYS

##data
```{r}

crest_surveys <- read_csv("./data/surveys_RAW.csv") %>% 
  dplyr::select(2:ncol(.))

#density data

crest_density <- crest_surveys %>% 
  mutate(prop_DC = (DC/(DC+LC)),
         prop_LC = (LC/(DC+LC)),
         dens_PM = (as.numeric(adult)+juv)/col_m2,
         dens_adult = as.numeric(adult)/col_m2)
  
  crest_summary <- crest_density %>% 
    group_by(type) %>% 
    summarise(n=length(dens_PM),
              n_transect = length(unique(Code)),
              mean_25 = mean(dens_PM[prop_DC<=0.25]),
              mean_75 = mean(dens_PM[prop_DC>=0.75]),
              SEM_25 = sd(dens_PM[prop_DC<=0.25])/sqrt(length(dens_PM[prop_DC<=0.25])),
              SEM_75 = sd(dens_PM[prop_DC>=0.75])/sqrt(length(dens_PM[prop_DC>=0.75])))

  #summary coral data
crest_transects <- crest_density %>% 
  dplyr::select(2, 8:18) %>% 
  distinct() %>% 
  mutate(LCC = LC/100,
         DCC = DC/100,
         prop_o_LC = replace_na(o_LC/(LC+DC),0),
         prop_o_DC = replace_na(o_DC/(DC+LC),0)) %>% 
  dplyr::select(9:10,12,15:16) %>% 
  pivot_longer(names_to = "type",values_to = "prop.occ", cols = 4:5 ) %>% 
  mutate(type = recode(type, "prop_o_DC" = "Dead",
                "prop_o_LC" = "Live")) 

#only occupied corals

occ_transects <- crest_density %>% 
  filter(adult>0) %>% 
  dplyr::group_by(Code, prop_DC) %>% 
  dplyr::mutate(ttl_occ = length(Code)) %>% 
  ungroup() %>% 
  dplyr::group_by(Code, type, prop_DC) %>% 
 dplyr::summarise(occ_freq = length(Code),
                  oc_prop = occ_freq/ttl_occ) %>% 
  distinct()

occ_summary <- occ_transects %>% 
    group_by(type) %>% 
    summarise(n=length(oc_prop))
  
```

#analysis
```{r}
dens.lm <- lm(log(dens_PM)~prop_DC*type, data = crest_density)

summary(dens.lm)

plot(dens.lm)

library(broom)
model.diag.metrics <- augment(dens.lm)
model.diag.metrics %>%
  top_n(3, wt = .cooksd)

```
#Q1 - DD HS?
**Is there evidence for DD habitat selection on natural reefs, and is this influenced by the health of the colony you're on?**  
 *YES* There is a significant relationship between the density of moluccensis on coral colonies and the proportion dead coral on a transect, and the slope of this relationship differs depending on whether the coral colony is live or dead (two-way anova, interaction term t=2.5, df(3,375), p=0.01). Adjusted R-squared:  0.1697
 

##figure
```{r}

#density @ colony level
(dens_adult <- ggplot(data =  crest_density, aes(x=prop_DC, y = dens_adult, fill = type,  group = type))+
  geom_point(size = 3, alpha = 0.5, shape = 21, aes(fill = type))+
   stat_smooth(method = "lm", colour="black", size = 0.5, alpha =0.3, show.legend = F)+ #linear
    # stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)+ #quadratic
    # stat_smooth(method = "loess")+ #even loess is straight, ha!
  scale_fill_manual(values = cols_)+
  scale_colour_manual(values = cols_)+
  labs(y = expression(Adult~density~m^-2), x = "")+
  scale_y_continuous(limits = c(0,450),  expand = c(0,20))+
  # facet_wrap(~period)+
  theme_DDHS+
  theme(legend.position = c(0.15,0.85),
        legend.title = element_blank()))

#prop occupied @ transect level - by all colonies
(prop_occupied <- ggplot(data = crest_transects, aes(x=prop_DC, y=prop.occ, fill= type,  group=type) )+
  geom_point(size = 3, alpha = 0.5, shape = 21, aes(fill = type))+
  stat_smooth(method = "lm",  size = 0.5, colour = "black", alpha = 0.3, show.legend = F)+
  scale_fill_manual(values = cols_)+
  scale_colour_manual(values = cols_)+
  labs(y = "Proportion occupied overall", x = "")+
   scale_y_continuous(expand = c(0,0.01))+
  theme_DDHS+
  theme(legend.position = "none"))

# fig_2 <- ggarrange(dens_adult, prop_occupied, nrow=1, align = "v", labels = c("a", "b"))
  # annotate_figure(fig_2, bottom = "Proportion dead coral (transect)")

# ggsave("./figures/Figure.2.jpg", height = 10, width = 20, units = "cm")
```
##change point analysis

```{r}
library(mcp)

#define a priori model
model = list(
  dens_adult ~ 1 + prop_DC,  #intercept & slope
   ~ 0 + prop_DC + I(prop_DC^2) #quadratic
)

model = list(
  dens_adult ~ 1 + prop_DC,  #intercept & slope
   ~ 0 + prop_DC  #different slope?
)

model = list(
  dens_adult ~ 1 + prop_DC,  #intercept & slope
   ~ sigma(0 + prop_DC)  #different variance?
)

fit = mcp::mcp(model, subset(crest_density, type=="Live"))

plot(fit, q_predict = T)+
  theme_classic()+
  labs(y = "density m2", x = "proportion dead coral (transect)")

# ggsave("./figures/S1.png", width = 15, height = 10, units = "cm" )

print(fit$prior)
summary(fit)
plot_pars(fit, regex_pars = "sigma_")
```

#Q2 - Ecological trap?
**Does the presence of adults influence the habitat choice of juveniles, and does this effect differ depending on the identity of the coral colony?**

Yes, juveniles are more likely to recruit to a dead coral if there are adults present, sig interaction p<0.0001, Adj R2 = 0.74, F-statistic: 138.4 on 4 and 183 DF,

##Transect level Manlys
###data

```{r}

tran_used <- crest_surveys %>% 
  mutate(adult.p = if_else(adult>0, "adult present", "adult absent"),
         ID = str_c(type, adult.p, sep="_")) %>% 
  group_by(Code, ID) %>% 
  summarise(freq=sum(juv2)) %>% 
  pivot_wider(names_from = ID, values_from = freq, values_fill = 0) 

crl.avail <- crest_surveys %>% 
  mutate(adult.p = if_else(adult>0, "adult present", "adult absent"),
         ID = str_c(type, adult.p, sep="_"),
         LC.prop = LC/(LC+DC),
         DC.prop = DC/(LC+DC)) %>% 
  dplyr::select(Code, LC.prop, DC.prop) 

  #calc 
juvs.avail <- tran_used%>%  
  full_join(crl.avail, by="Code") %>% 
  dplyr::mutate("Live_adult present" = LC.prop,
                "Live_adult absent" = LC.prop,
                "Dead_adult present" = DC.prop,
                "Dead_adult absent" = DC.prop) %>% 
  dplyr::select(1:5) %>%
  distinct() %>% 
   column_to_rownames(var = "Code")
  
tran_used <- tran_used %>% 
  column_to_rownames(var = "Code")

library(adehabitatHS)  
wiRatio <- widesIII(u=tran_used, a=juvs.avail, alpha =0.05)

plot(wiRatio)

#extract data
manly_data <- data.frame(wiRatio$wij) %>% 
  rownames_to_column(var = "Code") %>% 
  inner_join(crl.avail, by = "Code") %>% 
  pivot_longer(cols = 2:5, names_to = "type", values_to = "MSC") %>% 
  distinct() %>% 
  mutate(c.type = fct_relevel(str_split(type, "_", simplify = T)[,1], "Dead"),
         adult = str_split(type, "_", simplify = T)[,2]) 
  

```

###figure
```{r}
#plot it

(eco_trap <- ggplot(data = manly_data, aes(x=DC.prop, y= log(MSC), fill = c.type, shape = adult))+
  stat_smooth(aes(colour = c.type), method = "lm", se = T, size =0.5, show.legend = F)+
  geom_point(size = 3, alpha = 0.8, show.legend = T)+
  scale_shape_manual(values = c(21,22), name = "adults", labels = c("absent", "present"))+
  scale_fill_manual(values = cols_, name = "colony type")+
  geom_abline(intercept = 0, slope = 0, linetype = "dashed")+
  guides(fill = guide_legend(override.aes=list(shape=21)),
         colour = guide_none())+
  labs(x = "Proportion dead coral")+
  theme_classic()+
  annotate("text", label = "preference", x = 0.05, y=0.2)+
  annotate("text", label = "avoidance", x = 0.05, y=-0.2))

# ggsave("./figures/Figure 3_V2.jpg", width = 20, height = 15, units="cm")

```
### analysis
```{r}

manly_stat <- lm(log(MSC+1)~DC.prop+c.type*adult, data = manly_data)
summary(manly_stat)
plot(manly_stat)

```
##Q2b - Demographic consequence

**One of the assumptions of ecological trap is that there has to be a demonstrated negative demographic consequence of being trapped. So, what happens to the density of P. mol after 1 year?**

The density of P. mol adults declined by 54% (t1 0.39, t2 0.17), and juvenils by 17% (t1 0.06, t2 0.04), and this difference is significant for  adutls (adults F = 13.45 (1,87), P<0.001) , but not for juveniles (juveniles F = 0.7 (1,87), P=0.4).

###Data
```{r}

T2_raw <- read_csv("./data/surveys_RAW_T2.csv")

```
###analysis

```{r}
T2_dens <- T2_raw %>%
  dplyr::mutate(adult_den = adult/100,
                juv_den = juv/100) %>% 
  dplyr::mutate(period = "t2") %>% 
  dplyr::select("Code", "adult", "juv", "adult_den", "juv_den", "period")

diff_dens <- crest_surveys %>%
  group_by(Code) %>% 
 mutate( adult = replace_na(as.double(adult), 0)) %>% 
  summarise(adult = sum(adult),
            juv = sum(juv)) %>% 
  dplyr::mutate(adult_den = adult/100,
                juv_den = juv/100) %>% 
  dplyr::mutate(period = "t1") %>% 
  add_row(T2_dens)

#analysis
summary(aov(adult_den~period, data = diff_dens))
summary(aov(juv_den~period, data = diff_dens))

#summarise
T2_summary <- T2_raw %>%
  dplyr::mutate(adult_den = adult/100,
                juv_den = juv/100) %>% 
   summarise(mean_adult = mean(adult_den),
            SEM_adult = sd(adult_den/sqrt(length(adult_den))),
            mean_juv = mean(juv_den),
            SEM_juv = sd(juv_den/sqrt(length(juv_den)))) %>%
  dplyr::mutate(period = "t2") 

diff_summary <- crest_surveys %>%
  group_by(Code) %>% 
 mutate( adult = replace_na(as.double(adult), 0)) %>% 
  summarise(adult = sum(adult),
            juv = sum(juv)) %>% 
  dplyr::mutate(adult_den = adult/100,
                juv_den = juv/100) %>% 
  summarise(mean_adult = mean(adult_den),
            SEM_adult = sd(adult_den/sqrt(length(adult_den))),
            mean_juv = mean(juv_den),
            SEM_juv = sd(juv_den/sqrt(length(juv_den)))) %>%
  dplyr::mutate(period = "t1") %>% 
  add_row(T2_summary)
  
```



# Q3 - Experiment
Does DD HS occur? Is it influenced by the identity of the alternate coral colony, when there is an adult present ( Ie teasing apart cue of conspecific vs cue from live coral)?
**YES, The density of adults on the established coral colony had a significant effect on the habitat choice of P. moluccensis juveniles in the habitat selection experiment X2=62.7, df=2, p<0.0001**

###Data
```{r}
expt_data <- read_csv("./data/Choice.csv") %>% 
  janitor::clean_names() %>% 
  group_by(treatment, density, code) %>% 
  mutate(n.trials = length(trial)) %>% 
  mutate(treatment = recode(treatment, 
                            "1.Live" = "Live",
                            "2.Degraded" = "Dead"))

#for points only

expt_points <- expt_data %>% 
  dplyr::select(2:5) %>% 
  distinct()
```
###Analysis
```{r}
##Test
choice_log=glm(data=expt_data, code~treatment+density, family=binomial(logit))
summary(choice_log)
anova(choice_log, test="Chisq")


### Calculating likelihood ratio for appropriate model ####
with (choice_log, null.deviance-deviance)
with (choice_log, df.null-df.residual)
with (choice_log, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE))

### Walds Chisq values

wald.test(b = coef(choice_log), Sigma = vcov(choice_log), Terms = 2)
wald.test(b = coef(choice_log), Sigma = vcov(choice_log), Terms = 3)


##predicting LD50 (for p specific values)

choice_H=glm(data=subset(expt_data, expt_data$treatment=="Live"), code~density, family=binomial(logit)) #subset dataset for each separate model (ie outcomes)
predict_H=(dose.p(choice_H, p=c(0.5, 0.75,0.8, 0.9))) #calculates the "dose" (=y-value) where p (=x-value) equals the numbers you specify

choice_D=glm(data=subset(expt_data, expt_data$treatment=="Dead"), code~density, family=binomial(logit))
predict_D=(dose.p(choice_D, p=c(0.5, 0.75,0.8, 0.9)))

52.69085*1.96 #95%CI Degraded
35.98636*1.96 #95%CI Healthy
```

###Figure
```{r}
(expt_plot <- ggplot(data = expt_data, aes(x=density, y=code ))+
 geom_point(data = expt_points, aes(size = n.trials, fill = treatment), shape = 21,  alpha = 0.5)+
  stat_smooth(method="glm",method.args = list(family = "binomial"), se=T,  size=0.7, aes(fill=treatment, linetype = treatment, colour = treatment))+ 
   scale_fill_manual(values = cols_)+ 
   #LD50 line Live (use values here that you caclculated using dose.p)
geom_segment( aes(y = 0.5, yend = 0.5, xend = 316.8520, x = -Inf),size=0.5, linetype = "dotted") + 
geom_segment( aes(x=316.8520,y=-Inf,xend=316.8520,yend=0.5),size=0.5, linetype = "dotted") +

#LD50 line Dead (use values here that you caclculated using dose.p)
geom_segment( aes(y = 0.5, yend = 0.5, xend = 411.8904,x =316.8520 ),size=0.5,linetype = "dotted") +
geom_segment( aes(x=411.8904, xend=411.8904, yend=0.5,y = -Inf),size=0.5, linetype = "dotted") + 
  labs(x="Density on established colony", y="Colony choice")+
  guides(fill=guide_none(),
         linetype = guide_legend(title="Alternate \ncolony type"),
         colour = guide_legend(title="Alternate \ncolony type", override.aes=list(fill=NA)),
         size = guide_legend(title="Number of \ntrials") )+
  theme_DDHS)

# ggsave("./figures/Figure 4.svg", width = 15, height = 10, units = "cm")
```


